A Partitioning Algorithm for Maximum Common Subgraphs\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant numbers EP/K503058/1 and EP/M508056/1]}

### 2017-02-14 paper/paper.tex
Added a bit more to the comparison with other algorithms

\section{Comparison with Existing Algorithms}\label{sec:comparison}
The  \McSplit\ algorithm may be considered as a version of the
constraint programming, forward-checking (CP-FC) algorithm of \cite{DBLP:conf/cp/NdiayeS11}.
The vertices in $P$ correspond to variables in CP-FC and the 
vertices in $T$ to the domains of variables in CP-FC. A label class $\langle P,T \rangle$ then corresponds to a set 
of variables ($P$) in CP-FC  with identical domains ($T$). Where CP-FC uses a soft allDifferent constraint to filter and compute 
a bound, \McSplit\ computes a bound from each label class in the future, i.e. it performs Hall Set reasoning on the fly.
\iffalse
\McSplit\ takes $\BigO{n}$ time at each search node. Furthermore, \McSplit\ has a low start up cost. The clique encoding requires the
construction of the association graph prior to search

Our data structure lets us compute good heuristics cheaply, which is why we take
slightly fewer search nodes than CP-FC.

Clique gives us very strong propagation.

@@

%? Mention partition backtracking
\fi

@@

\paragraph{Directed graphs without edge labels} Let $A_G$ and $A_H$ be adjacency matrices
of $G$ and $H$. For each
vertex pair $(t,u)$ in graph $G$ or $H$, the adjacency matrix entry 
takes the value 0 if the vertices are
not adjacent, 1 if the two vertices share a single edge in the direction $t
\rightarrow u$, 2 if they share a single edge in the direction $u \rightarrow
t$, and 3 if there are edges in both directions. Where lines 14-19 of the
basic algorithm split the label class $\langle P',T' \rangle$ according to whether
vertices are adjacent to $v$ (resp. $w$), we now perform a four-way split where
each vertex is classified according to the label on its adjacency matrix from $v$ (resp. $w$). 
This is shown in Algorithm~\ref{labDirAlg}, where $L=\{0,1,2,3\}$.

\begin{algorithm}
\DontPrintSemicolon
\nl    \For {$l \in L$}{
\nl        \bf{Let} $P^{''} \gets \{ u \in P^{'} \mid u \neq v \wedge A_G[v][u] = l \}$ \;
\nl        \bf{Let} $T^{''} \gets \{ u \in T^{'} \mid u \neq w \wedge A_H[w][u] = l \}$ \;
\nl        \lIf {$P^{''} \neq \emptyset$ \bf{and} $T^{''} \neq \emptyset$}{$future^{'} \gets future^{'} + \langle P^{''} , T^{''} \rangle$}
       }
\caption{Replacement for lines 14--19 in directed in labelled cases}
\label{labDirAlg}
\end{algorithm}

@@

>>>>

\section{Comparison with Existing Algorithms}\label{sec:comparison}
The  \McSplit\ algorithm may be considered as a version of the
constraint programming, forward-checking (CP-FC) algorithm of \cite{DBLP:conf/cp/NdiayeS11}.
The vertices in $P$ correspond to variables in CP-FC and the 
vertices in $T$ to the domains of variables in CP-FC. A label class $\langle P,T \rangle$ then corresponds to a set 
of variables ($P$) in CP-FC  with identical domains ($T$). Where CP-FC uses a soft allDifferent constraint to filter and compute 
a bound, \McSplit\ computes a bound from each label class in the future, i.e. it performs Hall Set reasoning on the fly.

\McSplit\ takes $\BigO{n}$ time at each search node. Furthermore, \McSplit\ has a low start up cost. The clique encoding requires the
construction of the association graph prior to search

@@

? Mention partition backtracking

LARGELY FROM CIARAN'S EMAIL; NEEDS WORK:

\begin{algorithm}
\DontPrintSemicolon
\nl    \For {$q \in \N(b,G)$}{
\nl        remove $t$ from $D'_q$ \;
\nl        remove the inverse neighbourhood of $t$ from $D'_q$ \;
       }
\nl    \For {$q \in \N(b,\overline{G})$}{
\nl        remove $t$ from $D'_q$ \;
\nl        remove the neighbourhood of $t$ from $D'_q$ \;
       }
\caption{Replacement for \linerangeref{NewPWithEdge}{InnerLoopEnd} in CP algorithm}
\label{cpAlg}
\end{algorithm}

The first difference is in the underlying data structures: James uses a clever
partitioning data structure, not a conventional domain store. With the
algorithm described this way, I think the proof that domains are partitioned
into disjoint sets is almost immediate.

The second difference is in the bound function: the French code constructs a
maximum cardinality matching, with some special weighting trickery for $\bot$,
whilst James just sums the smaller of the two sides of each partition. If you
think a bit about what the matching looks like, you can see that these two give
the same answer, but at very different speeds.

The third difference is in the details of the variable ordering heuristic.

@@

\paragraph{Directed graphs without edge labels} Let $A_G$ and $A_H$ be
adjacency matrices of $G$ and $H$. For each vertex pair $(t,u)$ in graph $G$ or
$H$, the adjacency matrix entry takes the value 0 if the vertices are not
adjacent, 1 if the two vertices share a single edge in the direction $t
\rightarrow u$, 2 if they share a single edge in the direction $u \rightarrow
t$, and 3 if there are edges in both directions. Where
\linerangeref{NewPWithEdge}{InnerLoopEnd} of the basic algorithm split the
label class $\langle P',T' \rangle$ according to whether vertices are adjacent
to $v$ (resp. $w$), we now perform a four-way split where each vertex is
classified according to the label on its adjacency matrix from $v$ (resp. $w$).
This is shown in Algorithm~\ref{labDirAlg}, where $L=\{0,1,2,3\}$.

\begin{algorithm}
\DontPrintSemicolon
\nl    \For {$l \in L$}{
\nl        \bf{Let} $P^{''} \gets \{ u \in P^{'} \mid u \neq v \wedge A_G[v][u] = l \}$ \;
\nl        \bf{Let} $T^{''} \gets \{ u \in T^{'} \mid u \neq w \wedge A_H[w][u] = l \}$ \;
\nl        \lIf {$P^{''} \neq \emptyset$ \bf{and} $T^{''} \neq \emptyset$}{$future^{'} \gets future^{'} + \langle P^{''} , T^{''} \rangle$}
       }
\caption{Replacement for \linerangeref{NewPWithEdge}{InnerLoopEnd} in directed and labelled cases}
\label{labDirAlg}
\end{algorithm}

@@

? Portfolio algorithms---finding out at what level of labelledness clique becomes best etc.
